Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-07-10 patchy-selection-paper.tex
factor of 2

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in models of spatially varying selection.
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{s_m}$. 
Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panmictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate $p_e$ by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
More general work on the mathematically equivalent problem of density-dependent population growth 
has obtained much more general criteria
under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.

>>>>

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in models of spatially varying selection.
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{2s_m}$. 
Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panmictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate $p_e$ by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
More general work on the mathematically equivalent problem of density-dependent population growth 
has obtained much more general criteria
under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.
